NGO Repression as a Predictor of Worsening Human Rights Abuses

### 2021-11-19 README.Rmd
Better more stable project name

>>>>

### `r emoji::emoji_glue(":warning::bird:")`: "Cautioning canary" project name

Because project titles change all the time with revisions, rewriting, and peer review, we use "cautioning canary" as an internal-to-us project name that won't change. (We chose this name before the [**codename**](http://svmiller.com/codename/) was invented—had it existed already, we would have used that to generate the name.)

### 2021-11-19 manuscript/appendix.Rmd
Better more stable project name

---
title: 'Appendix for "NGO Repression as a Predictor of Worsening Human Rights Abuses"'
short-title: 'Appendix for "NGO Repression"'
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "November 19, 2021"
# published: Under review
code-repo: "Access the code and full analysis notebook at <https://stats.andrewheiss.com/canary-ngos/>"
reference-section-title: References
bibliography: bibliography.bib
toc: false
appendix: true
mainfont: Cochineal
sansfont: Inter
editor_options: 
  chunk_output_type: console
---

>>>>

---
title: 'Appendix for "NGO Repression as a Predictor of Worsening Human Rights Abuses"'
short-title: 'Appendix for "NGO Repression"'
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "November 19, 2021"
# published: Under review
code-repo: "Access the code and full analysis notebook at <https://stats.andrewheiss.com/cautioning-canary/>"
reference-section-title: References
bibliography: bibliography.bib
toc: false
appendix: true
mainfont: Cochineal
sansfont: Inter
editor_options: 
  chunk_output_type: console
---

### 2021-11-19 manuscript/manuscript.Rmd
Better more stable project name

---
title: "NGO Repression as a Predictor of Worsening Human Rights Abuses"
short-title: "NGO Repression and Human Rights"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "November 19, 2021"
published: "Conditionally accepted at the *Journal of Human Rights*"
code-repo: "Access the code and full analysis notebook at <https://stats.andrewheiss.com/canary-ngos/>"
abstract: >-
  An increasing number of countries have recently cracked down on non-governmental organizations (NGOs). Much of this crackdown is sanctioned by law and represents a bureaucratic form of repression that could indicate more severe human rights abuses in the future. This is especially the case for democracies, which unlike autocracies, may not aggressively attack civic space. We explore if crackdowns on NGOs predict broader human rights repression. Anti-NGO laws are among the most subtle means of repression and attract lesser domestic and international condemnation compared to the use of violence. Using original data on NGO repression, we test whether NGO crackdown is a predictor of political terror, and violations of physical integrity rights and civil liberties. We find that while de jure anti-NGO laws provide little information in predicting future repression, their patterns of implementation—or de facto civil society repression—predicts worsening respect for physical integrity rights and civil liberties.
# keywords: civil society, NGOs, repression, human rights, political terror, civil liberties
word-count: "8,343"
thanks: >-
  We would like to thank participants of the 2021 International Studies Association (ISA) convention, and the "Human Rights on the Edge: The Future of International Human Rights Law and Practice" workshop hosted by the Global Human Rights Hub at Arizona State University for their comments and feedback.
reference-section-title: References
bibliography: bibliography.bib
toc: false
title-page: true
endnotes: true
endfloat: false
mainfont: Cochineal
sansfont: Inter
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of International Affairs, Department of International Affairs, Lewis \& Clark College; Andrew Heiss, PhD, Assistant Professor of Public Management, Andrew Young School of Policy Studies, Georgia State University. 
correspondence: >-
  Correspondence concerning this article should be addressed to Andrew Heiss, PhD, Department of Public Management and Policy, Andrew Young School of Policy Studies, Georgia State University, 14 Marietta Street NW, Suite 357, Atlanta, GA 30303. Contact: aheiss@gsu.edu
editor_options: 
  chunk_output_type: console
---

@@

```{r correct-not-correct}
# These numbers come from the analysis notebook at 
# https://stats.andrewheiss.com/canary-ngos/04_predictions.html#Prediction_diagnostics
pts_pred_table <- tribble(
  ~Prediction, ~Baseline, ~total, ~advocacy, ~entry, ~funding, ~repression,
  "Wrong",     109,       107,    110,       108,    108,      111,
  "Correct",   376,       378,    375,       377,    376,      374
) %>% 
  rename("Total legal barriers included" = total,
         "Barriers to advocacy included" = advocacy,
         "Barriers to entry included" = entry,
         "Barriers to funding included" = funding,
         "Civil society repression included" = repression)

@@

```{r tbl-rmse}
# These numbers also come from the analysis notebook at 
# https://stats.andrewheiss.com/canary-ngos/04_predictions.html
baseline_rmse <- 0.1862156
tbl_rmse <- tribble(
  ~`Latent human rights values`, ~RMSE,
                  "Baseline",    baseline_rmse,
      "Total legal barriers",    0.1872181,
      "Barriers to advocacy",    0.1872011,
         "Barriers to entry",    0.1862635,
       "Barriers to funding",    0.1879768,
  "Civil society repression",    0.1857038
) %>% 
  mutate(`% change from baseline` = (RMSE - baseline_rmse) / baseline_rmse,
         `% change from baseline` = percent(`% change from baseline`, accuracy = 0.01)) %>% 
  mutate(`% change from baseline` = recode(`% change from baseline`, "0.00%" = "—")) %>% 
  mutate(RMSE = round(RMSE, 4))

>>>>

---
title: "NGO Repression as a Predictor of Worsening Human Rights Abuses"
short-title: "NGO Repression and Human Rights"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "November 19, 2021"
published: "Conditionally accepted at the *Journal of Human Rights*"
code-repo: "Access the code and full analysis notebook at <https://stats.andrewheiss.com/cautioning-canary/>"
abstract: >-
  An increasing number of countries have recently cracked down on non-governmental organizations (NGOs). Much of this crackdown is sanctioned by law and represents a bureaucratic form of repression that could indicate more severe human rights abuses in the future. This is especially the case for democracies, which unlike autocracies, may not aggressively attack civic space. We explore if crackdowns on NGOs predict broader human rights repression. Anti-NGO laws are among the most subtle means of repression and attract lesser domestic and international condemnation compared to the use of violence. Using original data on NGO repression, we test whether NGO crackdown is a predictor of political terror, and violations of physical integrity rights and civil liberties. We find that while de jure anti-NGO laws provide little information in predicting future repression, their patterns of implementation—or de facto civil society repression—predicts worsening respect for physical integrity rights and civil liberties.
# keywords: civil society, NGOs, repression, human rights, political terror, civil liberties
word-count: "8,343"
thanks: >-
  We would like to thank participants of the 2021 International Studies Association (ISA) convention, and the "Human Rights on the Edge: The Future of International Human Rights Law and Practice" workshop hosted by the Global Human Rights Hub at Arizona State University for their comments and feedback.
reference-section-title: References
bibliography: bibliography.bib
toc: false
title-page: true
endnotes: true
endfloat: false
mainfont: Cochineal
sansfont: Inter
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of International Affairs, Department of International Affairs, Lewis \& Clark College; Andrew Heiss, PhD, Assistant Professor of Public Management, Andrew Young School of Policy Studies, Georgia State University. 
correspondence: >-
  Correspondence concerning this article should be addressed to Andrew Heiss, PhD, Department of Public Management and Policy, Andrew Young School of Policy Studies, Georgia State University, 14 Marietta Street NW, Suite 357, Atlanta, GA 30303. Contact: aheiss@gsu.edu
editor_options: 
  chunk_output_type: console
---

@@

```{r correct-not-correct}
# These numbers come from the analysis notebook at 
# https://stats.andrewheiss.com/cautioning-canary/04_predictions.html#Prediction_diagnostics
pts_pred_table <- tribble(
  ~Prediction, ~Baseline, ~total, ~advocacy, ~entry, ~funding, ~repression,
  "Wrong",     109,       107,    110,       108,    108,      111,
  "Correct",   376,       378,    375,       377,    376,      374
) %>% 
  rename("Total legal barriers included" = total,
         "Barriers to advocacy included" = advocacy,
         "Barriers to entry included" = entry,
         "Barriers to funding included" = funding,
         "Civil society repression included" = repression)

@@

```{r tbl-rmse}
# These numbers also come from the analysis notebook at 
# https://stats.andrewheiss.com/cautioning-canary/04_predictions.html
baseline_rmse <- 0.1862156
tbl_rmse <- tribble(
  ~`Latent human rights values`, ~RMSE,
                  "Baseline",    baseline_rmse,
      "Total legal barriers",    0.1872181,
      "Barriers to advocacy",    0.1872011,
         "Barriers to entry",    0.1862635,
       "Barriers to funding",    0.1879768,
  "Civil society repression",    0.1857038
) %>% 
  mutate(`% change from baseline` = (RMSE - baseline_rmse) / baseline_rmse,
         `% change from baseline` = percent(`% change from baseline`, accuracy = 0.01)) %>% 
  mutate(`% change from baseline` = recode(`% change from baseline`, "0.00%" = "—")) %>% 
  mutate(RMSE = round(RMSE, 4))
